---
title: "566-hw4"
author: "Yumeng Gao"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

#HPC
## Problem 1: Make sure your code is nice
Rewrite the following R functions to make them faster. It is OK (and recommended) to take a look at Stackoverflow and Google

```{r}
m= matrix(1:16, ncol=4)
m
```


```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1(m)

fun1alt <- function(mat) {
  rowSums(mat)
}

fun1alt(m)

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2(m)

fun2alt <- function(mat) {
  t(apply(mat,1,cumsum))
  #ans= data.frame(mat)
  #cumsum(ans)
  #dt= as.data.table(mat)
  #x=cumsum(dt)
  #x
  #dt[, acc_sum := cumsum(count)]
}

fun2alt(m)

# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
t1= microbenchmark::microbenchmark(fun1(dat), fun1alt(dat))
print(t1, unit = "relative", check = "equivalent")


# Test for the second
t2= microbenchmark::microbenchmark(fun2(dat), fun2alt(dat))
print(t2, unit = "relative", check = "equivalent")
```

The last argument, check = “equivalent”, is included to make sure that the functions return the same result.


## Problem 2: Make things run faster with parallel computing
The following function allows simulating PI
```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

In order to get accurate estimates, we can run this function multiple times, with the following code:
```{r}
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

Rewrite the previous code using parLapply() to make it run faster. Make sure you set the seed using clusterSetRNGStream():
```{r}
# YOUR CODE HERE
system.time({
  # YOUR CODE HERE
  ans <- # YOUR CODE HERE
  print(mean(ans))
  # YOUR CODE HERE
})
```

#SQL
Setup a temporary database by running the following chunk
```{r}
# install.packages(c("RSQLite", "DBI"))

library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

When you write a new chunk, remember to replace the r with sql, connection=con. Some of these questions will reqruire you to use an inner join. Read more about them here https://www.w3schools.com/sql/sql_join_inner.asp

## Question 1
How many many movies is there avaliable in each rating catagory.

## Question 2
What is the average replacement cost and rental rate for each rating category.

## Question 3
Use table film_category together with film to find the how many films there are witth each category ID

## Question 4
Incorporate table category into the answer to the previous question to find the name of the most popular category.




























































