---
title: "566-hw2"
author: "Yumeng Gao"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

### Prepare the library
```{r}
library(dplyr)
library(tidyverse)
library(data.table)
library(R.utils)
library(lubridate)
library(leaflet)
library(webshot)
webshot::install_phantomjs()
```


# Data Wrangling

### Merge the individual and regional CHS datasets using the location variable.
```{r}

```


## 1.After merging the data, make sure you don’t have any duplicates by counting the number of rows. Make sure it matches.
```{r}

```

In the case of missing values, impute data using the average within the variables “male” and “hispanic”.
```{r}

```


## 2.Create a new categorical variable named “obesity_level” using the BMI measurement (underweight BMI<14; normal BMI 14-22; overweight BMI 22-24; obese BMI>24). To make sure the variable is rightly coded, create a summary table that contains the minimum BMI, maximum BMI, and the total number of observations per category.
```{r}

```


## 3.Create another categorical variable named “smoke_gas_exposure” that summarizes “Second Hand Smoke” and “Gas Stove.” The variable should have four categories in total.
```{r}

```


## 4.Create four summary tables showing the average (or proportion, if binary) and sd of “Forced expiratory volume in 1 second (ml)” and asthma indicator by town, sex, obesity level, and “smoke_gas_exposure.”
```{r}

```



# Looking at the Data (EDA)

> The primary questions of interest are: 
> 1. What is the association between BMI and FEV (forced expiratory volume)? 
> 2. What is the association between smoke and gas exposure and FEV? 
> 3. What is the association between PM2.5 exposure and FEV?

## 1.Facet plot showing scatterplots with regression lines of BMI vs FEV by “townname”.
```{r}

```


## 2.Stacked histograms of FEV by BMI category and FEV by smoke/gas exposure. Use different color schemes than the ggplot default.
```{r}

```


## 3.Barchart of BMI by smoke/gas exposure.
```{r}

```


## 4.Statistical summary graphs of FEV by BMI and FEV by smoke/gas exposure category.
```{r}

```


## 5.A leaflet map showing the concentrations of PM2.5 mass in each of the CHS communities.
```{r}

```


## 6.Choose a visualization to examine whether PM2.5 mass is associated with FEV.
```{r}

```









