---
title: "566-hw3"
author: "Yumeng Gao"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

> Prepare the library

```{r}
library(R.utils)
library(lubridate)
library(tidyverse)
library(tidytext)
library(dplyr)
library(dtplyr)
library(data.table)
library(tidyr)
library(ggplot2)
library(forcats)
library(stringr)
library(rvest)
library(xml2)
library(httr)
```


# APIs

## 1. Using the NCBI API, look for papers that show up under the term “sars-cov-2 trial vaccine.” Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r}
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")

counts <- as.character(counts)

stringr::str_extract(counts, "[0-9,]+")
```
There were totally 4006 papers on the website.


## 2. Using the list of pubmed ids you retrieved, download each papers’ details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r}
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db= "pubmed",
    term= "sars-cov-2 trial vaccine",
    retmax= 250
  ), 
)

# Extracting the content of the response of GET
ids <- httr::content(query_ids)
```

```{r}
# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "</?Id>")

head(ids)
```

Grab publications with pubmed id 
```{r}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db= "pubmed",
    id= paste(ids,collapse = ","),
    retmax= 250,
    rettype= "abstract"
    )
)

# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```


## 3. As we did in lab 7. Create a dataset containing the following:

* Pubmed ID number,
* Title of the paper,
* Name of the journal where it was published,
* Publication date, and
* Abstract of the paper (if any).

```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```

### Get the titles:
```{r}
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
titles[[1]]
```

```{r}
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
titles[[1]]
```

### Get the journals:
```{r}
journals <- str_extract(pub_char_list, "<Title>[[:print:][:space:]]+</Title>")
journals[[1]]
```

```{r}
journals <- str_remove_all(journals, "</?[[:alnum:]- =\"]+>")
journals[[1]]
```

### Get the publication dates:
```{r}
dates <- str_extract(pub_char_list, "<PubDate>[[:print:][:space:]]+</PubDate>")
dates[[1]]
```

```{r}
dates <- str_remove_all(dates, "</?[[:alnum:]- =\"]+>")
dates[[1]]
```

```{r}
dates <- str_replace_all(dates, "[[:space:]]+"," ")
dates[[1]]
```

### Get the abstracts:
```{r}
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
abstracts[[1]]
```

```{r}
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>") 
abstracts[[1]]
```

```{r}
abstracts <- str_replace_all(abstracts, "[[:space:]]+"," ")
abstracts[[1]]
```

Final dataset:
```{r}
database <- data.frame(
  PubMedId = ids,
  Title    = titles,
  Journal  = journals,
  PublicationDate= dates,
  Abstract = abstracts
)
knitr::kable(database[1:8,], caption = "Some papers about Sars-Cov-2 Trial Vaccine")
```



# Text Mining

## 1. Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?

The dataset contains 3241 abstracts from articles across 5 search terms.
```{r}
if (!file.exists("abstracts.csv")) {
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv", "abstracts.csv", method = "libcurl", timeout  = 60)
}
abs= read.csv("abstracts.csv")

str(abs)
table(abs$term)

abs= as_tibble(abs)
```

```{r}
abs %>%
  unnest_tokens(word, abstract) %>%
  count(word, sort = TRUE) %>%
  top_n(15, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

Remove stop words.
```{r}
abs %>%
  unnest_tokens(word, abstract) %>%
  anti_join(stop_words, by = c("word")) %>%
  count(word, sort = TRUE) %>%
  top_n(15, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

### For each term:
```{r}
abs %>%
  unnest_tokens(word, abstract) %>%
  group_by(term) %>%
  count(word, sort = TRUE) %>%
  filter( !(word %in% stop_words$word) & !grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(5, n) %>%
  arrange(term, desc(n)) %>%
knitr::kable()
```
1) covid

2) cystic fibrosis

3) meningitis

4) preeclampsia

5) prostate cancer


## 2. Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.

```{r}
abs %>%
  unnest_ngrams(bigram, abstract, n=2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(n, fct_reorder(bigram, n))) +
  geom_col()
```


## 3. Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the “document”) What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

```{r}
abs %>%
  unnest_tokens(word, abstract) %>%
  group_by(term) %>%
  count(word, sort = TRUE) %>%
  #filter(!(token %in% stop_words$word)) %>%
  top_n(5, n) %>%
  bind_tf_idf(word, term, n) %>%
  arrange(term, desc(tf_idf)) %>%
knitr::kable()
```



















