---
title: "566-hw3"
author: "Yumeng Gao"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

> Prepare the library

```{r}
library(R.utils)
library(lubridate)
library(tidyverse)
library(tidytext)
library(dplyr)
library(dtplyr)
library(data.table)
library(tidyr)
library(ggplot2)
library(forcats)
library(stringr)
library(rvest)
library(xml2)
library(httr)
```

# APIs

## 1. Using the NCBI API, look for papers that show up under the term "sars-cov-2 trial vaccine." Look for the data in the pubmed database, and then retrieve the details of the paper as shown in lab 7. How many papers were you able to find?

```{r}
website <- xml2::read_html("https://pubmed.ncbi.nlm.nih.gov/?term=sars-cov-2+trial+vaccine")

counts <- xml2::xml_find_first(website, "/html/body/main/div[9]/div[2]/div[2]/div[1]/div[1]")

counts <- as.character(counts)

stringr::str_extract(counts, "[0-9,]+")
```

There were totally 4009 papers on the website when searching for "sars-cov-2 trial vaccine".


## 2. Using the list of pubmed ids you retrieved, download each papers' details using the query parameter rettype = abstract. If you get more than 250 ids, just keep the first 250.

```{r}
query_ids <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/esearch.fcgi",
  query = list(
    db= "pubmed",
    term= "sars-cov-2 trial vaccine",
    retmax= 250
  ), 
)

# Extracting the content of the response of GET
ids <- httr::content(query_ids)
```

```{r}
# Turn the result into a character vector
ids <- as.character(ids)

# Find all the ids 
ids <- stringr::str_extract_all(ids, "<Id>[[:digit:]]+</Id>")[[1]]

# Remove all the leading and trailing <Id> </Id>. Make use of "|"
ids <- stringr::str_remove_all(ids, "</?Id>")

head(ids)
```

Grab first 250 publications with pubmed id
```{r}
publications <- GET(
  url   = "https://eutils.ncbi.nlm.nih.gov/entrez/eutils/efetch.fcgi",
  query = list(
    db= "pubmed",
    id= paste(ids,collapse = ","),
    retmax= 250,
    rettype= "abstract"
    )
)

# Turning the output into character vector
publications <- httr::content(publications)
publications_txt <- as.character(publications)
```


## 3. As we did in lab 7. Create a dataset containing the following:

-   Pubmed ID number,
-   Title of the paper,
-   Name of the journal where it was published,
-   Publication date, and
-   Abstract of the paper (if any).

```{r}
pub_char_list <- xml2::xml_children(publications)
pub_char_list <- sapply(pub_char_list, as.character)
```

### Get the titles:

```{r}
titles <- str_extract(pub_char_list, "<ArticleTitle>[[:print:][:space:]]+</ArticleTitle>")
titles[[1]]
```

```{r}
titles <- str_remove_all(titles, "</?[[:alnum:]- =\"]+>")
titles[[1]]
```

### Get the journals:

```{r}
journals <- str_extract(pub_char_list, "<Title>[[:print:][:space:]]+</Title>")
journals[[1]]
```

```{r}
journals <- str_remove_all(journals, "</?[[:alnum:]- =\"]+>")
journals[[1]]
```

### Get the publication dates:

```{r}
dates <- str_extract(pub_char_list, "<PubDate>[[:print:][:space:]]+</PubDate>")
dates[[1]]
```

```{r}
dates <- str_remove_all(dates, "</?[[:alnum:]- =\"]+>")
dates[[1]]
```

```{r}
dates <- str_replace_all(dates, "[[:space:]]+"," ")
dates[[1]]
```

### Get the abstracts:

```{r}
abstracts <- str_extract(pub_char_list, "<Abstract>[[:print:][:space:]]+</Abstract>")
abstracts[[1]]
```

```{r}
abstracts <- str_remove_all(abstracts, "</?[[:alnum:]- =\"]+>") 
abstracts[[1]]
```

```{r}
abstracts <- str_replace_all(abstracts, "[[:space:]]+"," ")
abstracts[[1]]
```

Final dataset:

```{r}
database <- data.frame(
  PubMedId = ids,
  Title    = titles,
  Journal  = journals,
  PublicationDate= dates,
  Abstract = abstracts
)
knitr::kable(database[1:5,], caption = "Some papers about Sars-Cov-2 Trial Vaccine")
```

# Text Mining

## 1. Tokenize the abstracts and count the number of each token. Do you see anything interesting? Does removing stop words change what tokens appear as the most frequent? What are the 5 most common tokens for each search term after removing stopwords?

The dataset contains 3241 abstracts from articles across 5 search terms.
```{r}
if (!file.exists("abstracts.csv")) {
  download.file(
    url = "https://raw.githubusercontent.com/USCbiostats/data-science-data/master/03_pubmed/pubmed.csv", "abstracts.csv", method = "libcurl", timeout  = 60)
}
abs= read.csv("abstracts.csv")

str(abs)
table(abs$term)

abs= as_tibble(abs)
```

```{r}
abs %>%
  unnest_tokens(word, abstract) %>%
  count(word, sort = TRUE) %>%
  top_n(15, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

For the top 15 common tokens, only "covid", "19", "patients", "cancer", and "prostate" are meaningful. Others are stop words and do not convey any interesting information.

### Remove stop words and numbers.

Since the only number we got above is "19" and it usually comes up with "covid" under this circumstance, we can filter it out.

```{r}
abs %>%
  unnest_tokens(word, abstract) %>%
  anti_join(stop_words, by = c("word")) %>%
  count(word, sort = TRUE) %>%
  filter( !grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(15, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

After removing stop words and numbers, the top 15 common tokens changed drastically. Now all those tokens have meanings and can help us to know these abstracts better. The most frequent word is "covid", indicating that most of these researches were about COVID-19.

### For each term:
```{r}
abs %>%
  unnest_tokens(word, abstract) %>%
  group_by(term) %>%
  count(word, sort = TRUE) %>%
  filter( !(word %in% stop_words$word) & !grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(5, n) %>%
  arrange(term, desc(n)) %>%
knitr::kable()
```

1)  covid: 
covid, patients, disease, pandemic, coronavirus / health (both appeared 647 times)

2)  cystic fibrosis:
fibrosis, cystic, cf, patients, disease

3)  meningitis:
patients, meningitis, meningeal, csf, clinical

4)  preeclampsia:
pre, eclampsia, preeclampsia, women, pregnancy

5)  prostate cancer:
cancer, prostate, patients, treatment, disease


## 2. Tokenize the abstracts into bigrams. Find the 10 most common bigram and visualize them with ggplot2.

```{r}
abs %>%
  unnest_ngrams(bigram, abstract, n=2) %>%
  count(bigram, sort = TRUE) %>%
  top_n(10, n) %>%
  ggplot(aes(n, fct_reorder(bigram, n))) +
  geom_col()
```

Here we can observe that most common bigrams were about COVID-19, Prostate cancer, and Preeclampsia.


## 3. Calculate the TF-IDF value for each word-search term combination. (here you want the search term to be the "document") What are the 5 tokens from each search term with the highest TF-IDF value? How are the results different from the answers you got in question 1?

```{r}
abs %>%
  unnest_tokens(word, abstract) %>%
  group_by(term) %>%
  count(word, term, sort = TRUE) %>%
  top_n(5, n) %>%
  bind_tf_idf(word, term, n) %>%
  arrange(term, desc(tf_idf)) %>%
knitr::kable()
```

There were lots of meaningless and redundant stop words in the table, so better get rid of them.

### Remove stop words and numbers:
```{r}
abs %>%
  unnest_tokens(word, abstract) %>%
  group_by(term) %>%
  count(word, term, sort = TRUE) %>%
  filter( !(word %in% stop_words$word) & !grepl(pattern = "^[0-9]+$", x = word)) %>%
  top_n(5, n) %>%
  bind_tf_idf(word, term, n) %>%
  arrange(term, desc(tf_idf)) %>%
knitr::kable()
```

1)  covid: 
covid, pandemic, coronavirus / health (both TF-IDF values equal 0.0826106), patients, disease

2)  cystic fibrosis:
fibrosis, cystic, cf, disease, patients

3)  meningitis:
meningitis, meningeal, csf, clinial, patients

4)  preeclampsia:
pre, eclampsia, preeclampsia, women, pregnancy

5)  prostate cancer:
cancer，prostate, treatment, disease, patients


The order of top 5 common tokens were changed for each term. TF-IDF not only focuses on the frequency of words but also provides the importance of the words. For example, although "patients" were appearing very frequently but its ranking here was lower than before for each term. Thus, the important tokens were ranked higher and the key information could be better presented by this method.







